Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-09-11 etm-transform/manuscript/main.tex
Section 4 needs extensive rewrite

\section{Application to Kyber}\label{sec:experimental-results}
CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism whose security is based on the conjecture hardness of the decisional Module Learning with Error problem. To achieve the the IND-CCA2 security, Kyber first constructs an IND-CPA secure public key encryption scheme based on \cite{lpr2013ideal}, then apply a slightly modified variation of the Fujisaki-Okamoto transformation described in \cite{hofheinz2017modular}. The resulting decapsulation routine is especially inefficient because Kyber's IND-CPA encryption routine incurs significantly more computational cost than the decryption routine. This makes Kyber a prime target for demonstrating the performance improvements enjoyed by the ``encrypt-then-MAC'' KEM construction.

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        % TODO: change to ML-KEM-512/768/1024
        ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
        \hline
        ML-KEM-512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
        \hline
        ML-KEM-512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
        \hline
        ML-KEM-512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
        \hline
        ML-KEM-512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
        \hline
        ML-KEM-768 & 192 bits & 129895 & 146405& 186445 \\
        \hline
        ML-KEM-768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
        \hline
        ML-KEM-768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
        \hline
        ML-KEM-768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
        \hline
        ML-KEM-768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
        \hline
        ML-KEM-1024 & 256 bits & 194921 & 199185 & 246245 \\
        \hline
        ML-KEM-1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
        \hline
        ML-KEM-1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
        \hline
        ML-KEM-1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
        \hline
        ML-KEM-1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
        \hline
        % TODO: also do a short comparison with Kyber Round 3
        % TODO: KeyGen, Encap, and Decap CPU cycles should be separate tables
        % TODO: also include MAC performances on different ciphertext sizes
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}

@@

\section{Experimental Evaluation}
We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available \cite{}. We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits. Our experiments were run on a desktop with a 2.3 GHz Intel Core i9 laptop (Coffee Lake) with 16 GB RAM. The codes were compiled using \texttt{clange 14.0.0}. 
%with \texttt{-std=c99 -O1 -fomit-frame-pointer flag}. 
Table~\ref{tab:ExpResults} reports the cycle counts for three different instances of Keyber, namely . The cycle counts reported in Table~\ref{tab:ExpResults} are the average of the cycle counts of 10000 executions of all algorithms. 
\begin{table}[h]
\centering
\begin{tabular}{|c|c||c|c| |c|c| |c|c|} \hline
\multicolumn{4}{|c|}{Kyber512}  & \multicolumn{2}{|c|}{Kyber512+Poly1305} & \multicolumn{2}{|c|}{Kyber512+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 1632 (or 32) & {\bf $\keygen$:} &  & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 800 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 768 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber768}  & \multicolumn{2}{|c|}{Kyber768+Poly1305} & \multicolumn{2}{|c|}{Kyber768+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 2400 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1184 & {\bf $\encap$:} & & {\bf $\encap^+$:} &  & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1088 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber1024}  & \multicolumn{2}{|c|}{Kyber1024+Poly1305} & \multicolumn{2}{|c|}{Kyber1024+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 3168 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1568 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1568 & {\bf $\decap$:} & & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\end{tabular}
\caption{Cycle counts in AVX2 on a Coffee Lake laptop}
\label{tab:ExpResults}
\end{table}

\begin{itemize}
\item Forward secrecy: Mutual authentication protocol using KEM, computing the number of key-excahnge per second for both Kyber and ours. 
\item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
\item Look at OpenSSL for GMAC with no AES-256. 
\item Also setup networking experiment on AWS
\end{itemize}

>>>>

\section{Application to Kyber}\label{sec:experimental-results}
% TODO: organize into three subsection:
% 1. Choosing a MAC
% 2. Comparing CPU cycles
% 3. Comparing KEX RTT

CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism whose security is based on the conjecture hardness of the decisional Module Learning with Error problem. To achieve the the IND-CCA2 security, Kyber first constructs an IND-CPA secure public key encryption scheme based on \cite{lpr2013ideal}, then apply a slightly modified variation of the Fujisaki-Okamoto transformation described in \cite{hofheinz2017modular}. The resulting decapsulation routine is especially inefficient because Kyber's IND-CPA encryption routine incurs significantly more computational cost than the decryption routine. This makes Kyber a prime target for demonstrating the performance improvements enjoyed by the ``encrypt-then-MAC'' KEM construction.

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
        \hline
        ML-KEM-512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
        \hline
        ML-KEM-512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
        \hline
        ML-KEM-512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
        \hline
        ML-KEM-512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
        \hline
        ML-KEM-768 & 192 bits & 129895 & 146405& 186445 \\
        \hline
        ML-KEM-768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
        \hline
        ML-KEM-768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
        \hline
        ML-KEM-768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
        \hline
        ML-KEM-768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
        \hline
        ML-KEM-1024 & 256 bits & 194921 & 199185 & 246245 \\
        \hline
        ML-KEM-1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
        \hline
        ML-KEM-1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
        \hline
        ML-KEM-1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
        \hline
        ML-KEM-1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
        \hline
        % TODO: also do a short comparison with Kyber Round 3
        % TODO: KeyGen, Encap, and Decap CPU cycles should be separate tables
        % TODO: also include MAC performances on different ciphertext sizes
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}

@@

\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel. \cite{bos2018crystals} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE). We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.

For clarity, we denote the party who sends the first message to be the client and the other party to be the server. Round trip time (RTT) is defined to be the time interval between the moment before the client starts generating ephemeral keypairs and the moment after the client derives the final session key. All experiements are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region. For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond) recorded.

\subsubsection{Unauthenticated key exchange (KE)}\label{sec:unauthenticated-key-exchange}
In unauthenticated key exchange, a single pair of ephemeral keypair $(\pk_e, \sk_e) \leftsample \keygen()$ is generated by the client. The client transmits the ephemeral public key $\pk_e$ to the server, who runs the encapsulation routine $(c_e, K_e) \leftsample \encap(\pk_e)$ and transmits the ciphertext $c_e$ back to the client. The client finally decapsulates the ciphertext to recover the shared secret $K_e \leftarrow \decap(\sk_e, c_e)$. The specific steps are described in algorithms \ref{alg:ke-client}, \ref{alg:ke-server}.

Note that in our implementation, a key derivation function (\texttt{KDF}) is applied to the ephemeral shared secret to derive the final session key. This step is added to maintain consistency with other authenticated key exchange protocols, where the final session key is derived from multiple shared secrets. The key derivation function is instantiated using Shake256, and the final session key is 256 bits in length.

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{KE}_\texttt{C}$}\label{alg:ke-client}
                \begin{algorithmic}[1]
                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
                    \State $\texttt{send}(\pk_e)$
                    \State $c_e \leftarrow \texttt{read}()$
                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
                    \State $K \leftarrow \texttt{KDF}(K)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{KE Client}}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{KE}_\texttt{S}$}\label{alg:ke-server}
                \begin{algorithmic}[1]
                    \State $\pk_e \leftarrow \texttt{read}()$
                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                    \State $\texttt{send}(c_e)$
                    \State $K \leftarrow \texttt{KDF}(K_e)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{KE Server}}
    \end{floatrow}
\end{figure}

The RTT comparison is summarized in table \ref{tbl:ke-rtt}

\begin{table}[H]
    \caption{Unauthenticated key exchange RTT comparison}\label{tbl:ke-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 92 & 97 \\
        \hline
        ML-KEM-512 + Poly1305 & 70 & 72 \\
        \hline
        ML-KEM-512 + GMAC & 73 & 76 \\
        \hline
        ML-KEM-512 + CMAC & 75 & 79 \\
        \hline
        ML-KEM-512 + KMAC & 76 & 78 \\
        \hline 
        ML-KEM-768 & 135 & 140 \\
        \hline
        ML-KEM-768 + Poly1305 & 99 & 104 \\
        \hline
        ML-KEM-768 + GMAC & 101 & 105 \\
        \hline
        ML-KEM-768 + CMAC & 103 & 109 \\
        \hline
        ML-KEM-768 + KMAC & 103 & 107 \\
        \hline
        ML-KEM-1024 & 193 & 199 \\
        \hline
        ML-KEM-1024 + Poly1305 & 138 & 141 \\
        \hline
        ML-KEM-1024 + GMAC & 140 & 145 \\
        \hline
        ML-KEM-1024 + CMAC & 143 & 148 \\
        \hline
        ML-KEM-1024 + KMAC & 144 & 149 \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Unilaterally authenticated key exchange (UAKE)}\label{sec:uakex}
In unilaterally authenticated key exchange, the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key. In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by demonstrating possession of the corresponding long-term secret key $\sk_S$. Details are described in algorithms \ref{alg:uake-client} and \ref{alg:uake-server}.

In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key. This helps the key exchange to achieve weak forward secrecy (citation needed).

Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes. KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage. (citation needed).

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{UAKE}_\texttt{C}(\pk_S)$}\label{alg:uake-client}
                \begin{algorithmic}[1]
                    \Require Server's long-term $\pk_S$
                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
                    \State $(c_S, K_S) \leftsample \encap(\pk_S)$
                    \State $\texttt{send}(\pk_e, c_S)$
                    \State $c_e \leftarrow \texttt{read}()$
                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{UAKE Client}}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{UAKE}_\texttt{S}(\sk_S)$}\label{alg:uake-server}
                \begin{algorithmic}[1]
                    \Require Server's long-term $\sk_S$
                    \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
                    \State $K_S \leftarrow \decap(\sk_S, c_S)$
                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                    \State $\texttt{send}(c_e)$
                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{UAKE Server}}
    \end{floatrow}
\end{figure}

\begin{table}[H]
    \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 145 & 151 \\
        \hline
        ML-KEM-512 + Poly1305 & 103 & 106 \\
        \hline
        ML-KEM-512 + GMAC & 106 & 110 \\
        \hline
        ML-KEM-512 + CMAC & 108 & 112 \\
        \hline
        ML-KEM-512 + KMAC & 109 & 113 \\
        \hline 
        ML-KEM-768 & 215 & 222 \\
        \hline
        ML-KEM-768 + Poly1305 & 144 & 150 \\
        \hline
        ML-KEM-768 + GMAC & 149 & 156 \\
        \hline
        ML-KEM-768 + CMAC & 153 & 160 \\
        \hline
        ML-KEM-768 + KMAC & 154 & 159 \\
        \hline
        ML-KEM-1024 & 310 & 318 \\
        \hline
        ML-KEM-1024 + Poly1305 & 202 & 209 \\
        \hline
        ML-KEM-1024 + GMAC & 212 & 228 \\
        \hline
        ML-KEM-1024 + CMAC & 212 & 218 \\
        \hline
        ML-KEM-1024 + KMAC & 213 & 220 \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Mutually authenticated key exchange (AKE)}\label{sec:akex}
Mutually authenticated key exchange is largely identical to unilaterally authenticated key exchange, except for that client authentication is required. This means that client possesses server's long-term public key and its own long-term secret key, while the server possesses client's long-term public key and its own long-term secret key. The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.

\begin{figure}[H]
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{AKE}_\texttt{C}(\pk_S, \sk_C)$}\label{alg:ake-client}
                \begin{algorithmic}[1]
                    \Require Server's long-term $\pk_S$
                    \Require Client's long-term $\sk_C$
                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
                    \State $(c_S, K_S) \leftsample \encap(\pk_S)$
                    \State $\texttt{send}(\pk_e, c_S)$
                    \State $(c_e, c_C) \leftarrow \texttt{read}()$
                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
                    \State $K_C \leftarrow \decap(\sk_e, c_C)$
                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{AKE Client}}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\texttt{AKE}_\texttt{S}(\sk_S, \pk_C)$}\label{alg:akex-server}
                \begin{algorithmic}[1]
                    \Require Server's long-term $\sk_S$
                    \Require Client's long-term $\pk_C$
                    \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
                    \State $K_S \leftarrow \decap(\sk_S, c_S)$
                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                    \State $(c_C, K_C) \leftsample \encap(\pk_C)$
                    \State $\texttt{send}(c_e, c_C)$
                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        }{\caption{AKE Server}}
    \end{floatrow}
\end{figure}

\begin{table}[H]
    \caption{AKE RTT comparison}\label{tbl:ake-rtt}
    \begin{tabular}{|c|c|c|}
        \hline
        KEM & median RTT ($\us$) & average RTT ($\us$) \\
        \hline
        ML-KEM-512 & 200 & 213 \\
        \hline
        ML-KEM-512 + Poly1305 & 133 & 138 \\
        \hline
        ML-KEM-512 + GMAC & 139 & 143 \\
        \hline
        ML-KEM-512 + CMAC & 143 & 148 \\
        \hline
        ML-KEM-512 + KMAC & 145 & 151 \\
        \hline 
        ML-KEM-768 & 294 & 301 \\
        \hline
        ML-KEM-768 + Poly1305 & 190 & 196 \\
        \hline
        ML-KEM-768 + GMAC & 197 & 210 \\
        \hline
        ML-KEM-768 + CMAC & 202 & 208 \\
        \hline
        ML-KEM-768 + KMAC & 204 & 210 \\
        \hline
        ML-KEM-1024 & 512 & 511 \\
        \hline
        ML-KEM-1024 + Poly1305 & 266 & 273 \\
        \hline
        ML-KEM-1024 + GMAC & 273 & 282 \\
        \hline
        ML-KEM-1024 + CMAC & 280 & 287 \\
        \hline
        ML-KEM-1024 + KMAC & 282 & 288 \\
        \hline
    \end{tabular}
\end{table}

\section{Experimental Evaluation}
We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available. We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits. Our experiments were run on a desktop with a 2.3 GHz Intel Core i9 laptop (Coffee Lake) with 16 GB RAM. The codes were compiled using \texttt{clange 14.0.0}. 
%with \texttt{-std=c99 -O1 -fomit-frame-pointer flag}. 
Table~\ref{tab:ExpResults} reports the cycle counts for three different instances of Keyber, namely . The cycle counts reported in Table~\ref{tab:ExpResults} are the average of the cycle counts of 10000 executions of all algorithms. 
\begin{table}[h]
\centering
\begin{tabular}{|c|c||c|c| |c|c| |c|c|} \hline
\multicolumn{4}{|c|}{Kyber512}  & \multicolumn{2}{|c|}{Kyber512+Poly1305} & \multicolumn{2}{|c|}{Kyber512+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 1632 (or 32) & {\bf $\keygen$:} &  & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 800 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 768 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber768}  & \multicolumn{2}{|c|}{Kyber768+Poly1305} & \multicolumn{2}{|c|}{Kyber768+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 2400 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1184 & {\bf $\encap$:} & & {\bf $\encap^+$:} &  & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1088 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber1024}  & \multicolumn{2}{|c|}{Kyber1024+Poly1305} & \multicolumn{2}{|c|}{Kyber1024+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 3168 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1568 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1568 & {\bf $\decap$:} & & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\end{tabular}
\caption{Cycle counts in AVX2 on a Coffee Lake laptop}
\label{tab:ExpResults}
\end{table}

\begin{itemize}
\item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
\item Look at OpenSSL for GMAC with no AES-256. 
\item Also setup networking experiment on AWS
\end{itemize}
